---
title: "hw9"
author: "Camille Okonkwo"
date: "2024-04-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Import Data and Preparation
```{r}
hwdata4 = read_csv("data/hwdata4.csv")

# data prep
hwdata4$treat = as.factor(hwdata4$treat)
hwdata4$tree = as.factor(hwdata4$tree)
```


1. Draw the spaghetti plots of the size of the tree on the growing time with separate panels for the two growing environments.
```{r}

```


2.	Write a random intercept model with the covariates including time, environment, and their interaction.
```{r}

```

3.	Was the trajectory of growth in the tree size over time different between the two environments? Perform a statistical test. Show the hypotheses, test statistic, p-value, and conclusion.
```{r}

```


4.	Calculate the within-subject correlation coefficient. What’s your observation? 
```{r}

```


5.	Write a random intercept and slope model with the covariates including time, environment, and their interaction, and random slope for time. 
```{r}

```

6.	Do we need the random intercept and slope model in addition to the random intercept model? Perform a statistical test. Show the hypotheses, test statistic, p-value, and conclusion
```{r}

```

