---
title: "hw9"
author: "Camille Okonkwo"
date: "2024-04-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Import Data and Preparation
```{r}
hwdata4 = read_csv("data/hwdata4.csv")

# data prep
hwdata4$treat = as.factor(hwdata4$treat)
hwdata4$tree = as.factor(hwdata4$tree)

hwdata4$treat2 = relevel(factor(hwdata4$treat,
                                labels =c("Ozone: grown under ozone environment", "Control: ozone free")),
                         ref = "Control: ozone free")
```


1. Draw the spaghetti plots of the size of the tree on the growing time with separate panels for the two growing environments.
```{r}
# spaghetti plot of size by time
hwdata4 |> 
  ggplot(aes(x = time, y = size, group = tree))+
  geom_point() +
  geom_line() +
  theme_bw() +
  stat_smooth(aes(group = 1),method = "lm",
              geom = "smooth",se =FALSE) +
  facet_grid(.~treat2, scales = "free") +
  labs(x = "Days after January 1st", y = "Size of the tree (log(height × diameter^2))") +
  ggtitle("Spaghetti Plot of Size of Spruce Trees Over Time by Treatment Environment")
```

3.	Was the trajectory of growth in the tree size over time different between the two environments? Perform a statistical test. Show the hypotheses, test statistic, p-value, and conclusion.
```{r}
library(nlme)
# wald test
hwdata4$treat = relevel(hwdata4$treat,
                        ref="ozone")

fit1 = lme(size ~ time + treat + time:treat,
           data = hwdata4, method = "ML",
           random = ~1 | tree,
           na.action = na.omit)
summary(fit1)

G = (0.000080/0.00017124)^2
G

pchisq(G, df = 1, lower.tail = F)
```

4.	Calculate the within-subject correlation coefficient. What’s your observation? 
```{r}
VarCorr(fit1)
varests = as.numeric(VarCorr(fit1))
rho = varests[1]/(varests[1]+varests[2])
rho
```

6.	Do we need the random intercept and slope model in addition to the random intercept model? Perform a statistical test. Show the hypotheses, test statistic, p-value, and conclusion
```{r}
fit2 = lme(size ~ time + treat + time:treat2,
           data = hwdata4,
           method = "ML",
           random = ~1 + time|tree)

summary(fit2)

logLik(fit2)
logLik(fit1)

G_2 = -2*(logLik(fit1)-logLik(fit2))
G_2

anova(fit1, fit2)

pchisq(G_2, df = 2, lower.tail = F)
```